---
title: "URSUSmerge"
format: html
editor: visual
---

## Merging Annual URSUS data

Authors: Taylor Domingos Date Began: 26 May 2021

```{r}
setwd("~/Dropbox/URSUS")
install.packages("janitor")
library (janitor)
library (plyr)
```

## SET UP: Read in Data Files

You will need to download both the Civilian-Officer and Incident CSV files from the OpenJustice California. Here is the link: https://openjustice.doj.ca.gov/data. Scroll to URSUS and you will see downloadable, separate files for each year. Download 2016-2021. Again there should be two CSV files for every year (Civilian-Officer and Incident CSV files).

```{r}
#2016
ursus2016 <- read.csv("URSUS_Civilian-Officer_2016.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
incident2016 <- read.csv("URSUS_Incident_2016.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)

#2017
ursus2017 <- read.csv("URSUS_Civilian-Officer_2017.csv", header = TRUE, stringsAsFactors = FALSE)
incident2017 <- read.csv("URSUS_Incident_2017.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)

#2018
ursus2018 <- read.csv("URSUS_Civilian-Officer_2018.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
incident2018 <- read.csv("URSUS_Incident_2018.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)

#2019
ursus2019 <- read.csv("URSUS_Civilian-Officer_2019.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
incident2019 <- read.csv("URSUS_Incident_2019.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)

#2020
ursus2020 <- read.csv("URSUS_Civilian-Officer_2020.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
incident2020 <- read.csv("URSUS_Incident_2020.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)

#2021
ursus2021 <- read.csv("URSUS_Civilian-Officer_2021.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)
incident2021 <- read.csv("URSUS_Incident_2021.csv", header = TRUE, sep = ",", stringsAsFactors = FALSE)

```

## CLEAN DATA: Convert all data to lowercase characters, trim whitespace, make sure all columns match across years

```{r}
# Data cleaning: trimming whitespace, lowercase column names
clean_data <- function(df) {
  df <- data.frame(tolower(trimws(as.matrix(df))), stringsAsFactors = FALSE)
  df <- clean_names(df, case = "snake")
  return(df)
}

# List of data frames
ursus_list <- list(ursus2016, ursus2017, ursus2018, ursus2019, ursus2020, ursus2021)
incident_list <- list(incident2016, incident2017, incident2018, incident2019, incident2020, incident2021)

# Apply the clean_data function to the lists
ursus_list <- lapply(ursus_list, clean_data)
incident_list <- lapply(incident_list, clean_data)

# Assign the cleaned data frames back to the original variables
for (i in seq_along(ursus_list)) {
  assign(paste0("ursus", 2016 + i - 1), ursus_list[[i]])
}

for (i in seq_along(incident_list)) {
  assign(paste0("incident", 2016 + i - 1), incident_list[[i]])
}

# Rename incident id column in 2019 data
names (ursus2019) [names(ursus2019) == "i_incident_id"] <- "incident_id"
names (incident2019) [names(incident2019) == "i_incident_id"] <- "incident_id"

# Add missing columns to all years
# Primary agency indicator is only included in 2017
incident2016$primary_agency_indicator <- NA
incident2018$primary_agency_indicator <- NA
incident2019$primary_agency_indicator <- NA
incident2020$primary_agency_indicator <- NA
incident2021$primary_agency_indicator <- NA

# Number of agencies involved is missing in 2016 and 2017
incident2016$num_involved_agencies <- NA
incident2017$num_involved_agencies <- NA

# Civilian crime qualifier is missing for 2018-2021
ursus2018$civilian_crime_qualifier <- NA
ursus2019$civilian_crime_qualifier <- NA
ursus2020$civilian_crime_qualifier <- NA
ursus2021$civilian_crime_qualifier <- NA

# Civilian_mental_status (2016-2017) is renamed signs civilian_signs_impairment_disability(2018-2021)
names (ursus2016) [names(ursus2016) == "civilian_mental_status"] <- "civilian_signs_impairment_disability"
names (ursus2017) [names(ursus2017) == "civilian_mental_status"] <- "civilian_signs_impairment_disability"

# Hi_islander_race missing 2016-2017
ursus2016$hi_islander_race <- NA
ursus2017$hi_islander_race <- NA


```

## MERGE DATA: Merge incident and ursus data and append years into one dataframe

```{r}
# Merge incident and civilian-officer files
data2016 <- merge (ursus2016, incident2016, by.x = "incident_id", by.y = "incident_id", all.x = TRUE, all.y = TRUE)
data2017 <- merge (ursus2017, incident2017, by.x = "incident_id", by.y = "incident_id", all.x = TRUE, all.y = TRUE)
data2018 <- merge (ursus2018, incident2018, by.x = "incident_id", by.y = "incident_id", all.x = TRUE, all.y = TRUE)
data2019 <- merge (ursus2019, incident2019, by.x = "incident_id", by.y = "incident_id", all.x = TRUE, all.y = TRUE)
data2020 <- merge (ursus2020, incident2020, by.x = "incident_id", by.y = "incident_id", all.x = TRUE, all.y = TRUE)
data2021 <- merge (ursus2021, incident2021, by.x = "incident_id", by.y = "incident_id", all.x = TRUE, all.y = TRUE)

# Append years into one dataframe
ursusmerged <- rbind (data2016,data2017,data2018,data2019,data2020,data2021)

# Write CSV
write.csv(ursusmerged, file = "ursusmerged1621.csv")
```
